The promise of community-driven preprints in ecology and evolution

### 2024-10-11 ms/ms.qmd
Dealing with a whole bunch of new comments

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. Preprints in biology have had a slower uptake relative to other disciplines [@Berg2016], but new discipline-specific preprint servers, such as *EcoEvoRxiv* (https://ecoevorxiv.org), provide a means by which ecologists and evolutionary biologists can disseminate research findings. Preprints attempt to break down barriers to scientific publishing by: 1) increasing the visibility of research and the speed at which research findings become available, which can lead to more citations [e.g., @colavizza2024analysis; @fu2019releasing]; 2) helping establish the precedence of research findings; 3) removing financial barriers to open access publication; and 4) enabling feedback from the research community [@Proulx2013; @bourne2017ten; @vale2015accelerating]. Ultimately, preprints can facilitate the rapid sharing of scientific knowledge that can have significant impacts on fundamental and applied knowledge globally [@ni2024preprint].

@@

### (b) *Cautious 'open'-mindedness of research in preprints*
In addition to speeding up dissemination, preprints and postprints can also be a useful way to ensure that research remains open and accessible to the research community irrespective of the accessibility of the final peer-reviewed paper [@bourne2017ten; @vale2015accelerating]. We evaluated whether preprints and postprints hosted at *EcoEvoRxiv*, and that were also published in a journal, were published open access. The open access status of each published article was obtained using the R package *roadoi* (v.`r utils::packageVersion("roadoi")`) to connect to the Unpaywall platform [@jahn2024]. Most of the published preprints and postprints were open access [`r sum_is_oa$valid_percent[2]*100`% (*n* = `r sum_is_oa$n[2]` out of `r sum(sum_is_oa$n[1:2])` where the status was known); [@fig-pubsummary]A]; however, `r sum_is_oa$valid_percent[1]*100`% (*n* = `r sum_is_oa$n[1]`) were published behind paywalls. For preprints and postprints published in open access journals, the type of open access also varied widely (e.g., Gold, Hybrid, Green OA etc., [@fig-pubsummary]B). Such a result may not be too surprising given that authors using preprint servers are probably already ‘pro-open access’; particularly given that *EcoEvoRxiv* is run by Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE).

Data and code sharing are also key components of open science [@roche2015public]. In the spirit of 'openness', we expected data and code sharing among preprints and postprints to be greater than in many papers published in research journals [@roche2015public; @o2021preferred]. Despite this, we found that `r (preprint_data[1,"n"]/sum(preprint_data[c(1,3),"n"]))*100`%  (*n* = `r preprint_data[1,2]`) of articles relying on data on *EcoEvoRxiv* did not share data, and `r (preprint_code[1,"n"]/sum(preprint_code[c(1,3),"n"]))*100`% (*n* = `r preprint_code[1,2]`) did not share code (counting only data-based articles, i.e., excluding reviews, commentaries or theoretical works).

Authors may be reluctant to share data and code for preprints because of the perceived concern that others may acquire and use their data and code before publication in a journal. Authors of `r pre_article_data[3,"percent"]*100`% (*n* = `r pre_article_data[3,2]`) of articles that did not share data at the preprint stage did ultimately share data when the article was published; whereas authors of `r pre_article_data[2,"percent"]*100`% (*n* = `r pre_article_data[2,2]`) never shared data. However, `r pre_article_data[1,"percent"]*100`% (*n* = `r pre_article_data[1,2]`) shared data at both stages. The same was true for code. Overall, `r pre_article_code[3,"percent"]*100`% (*n* = `r pre_article_code[3,2]`) preprints had no open code at the preprint stage but did at the published article stage and authors of `r pre_article_code[2,"percent"]*100`% (*n* = `r pre_article_code[2,2]`) preprints did not share code at either stage. However, `r pre_article_code[1,"percent"]*100`% (*n* = `r pre_article_code[1,2]`) shared code at both stages. Relatively low code and data-sharing practices in our sample is consistent with analyses of sharing practices for published articles (e.g., [@o2021preferred]), even for journals with strict public data archiving policies [@roche2015public].

@@

Our analysis has allowed us to better understand preprinting/postprinting practices in *EcoEvoRxiv*. Overall, *EcoEvoRxiv* articles are diverse but with primary research articles on vertebrates comprising most of the articles posted. North America, Europe and Australia use *EcoEvoRxiv* the most with very few non-English language articles to date. Submitting authors who were earlier in their career and more often with ‘male-associated names’ tended to use  *EcoEvoRxiv* the most. Articles posted to *EcoEvoRxiv* tend to take up to 8 months to become published with many articles not being open access. Code and data sharing was also relatively uncommon at the preprint stage. At the same time, we attempted to collect data on community discussion around preprints no such data was found on preprint landing pages, likely reflecting inadequate functionality and cross-linking with sources where such discussion is occurring. Based on the insights from our analysis, we provide recommendations to authors and the scientific community on ways they can further promote open and transparent research through preprints:

- First, share your data and code at the preprint stage. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings [@sarabipour2019value]. Reassuringly, sharing data and code is rarely associated with the 'scooping' of research findings [@soeharjono2021reported]. If authors are worried about data being used unintentionally, clear information surrounding its reuse can be included in a license (see https://choosealicense.com). Data can also be archived with an embargo on its reuse [@roche2014troubleshooting]. 

>>>>

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. Preprints in biology have had a slower uptake relative to other disciplines [@Berg2016], but new discipline-specific preprint servers, such as *EcoEvoRxiv* (https://ecoevorxiv.org), provide a means by which ecologists and evolutionary biologists can disseminate research findings. Preprints attempt to break down barriers to scientific publishing by: 1) increasing the visibility of research and the speed at which research findings become available, which can lead to more citations [e.g., @colavizza2024analysis; @fu2019releasing]; 2) helping establish the precedence of research findings; 3) removing financial barriers to open access publication; and 4) enabling earlier feedback from the research community [@Proulx2013; @bourne2017ten; @vale2015accelerating]. Ultimately, preprints can facilitate the rapid sharing of scientific knowledge that can have significant impacts on fundamental and applied knowledge globally [@ni2024preprint].

@@

### (b) *Cautious 'open'-mindedness of research in preprints*
In addition to speeding up dissemination, preprints and postprints can also be a useful way to ensure that research remains open and accessible to the research community irrespective of the accessibility of the final peer-reviewed paper [@bourne2017ten; @vale2015accelerating]. We evaluated whether preprints and postprints hosted at *EcoEvoRxiv*, and that were also published in a journal, were published open access. The open access status of each published article was obtained using the R package *roadoi* (v.`r utils::packageVersion("roadoi")`) to connect to the Unpaywall platform [@jahn2024]. Most of the published preprints and postprints were open access [`r sum_is_oa$valid_percent[2]*100`% (*n* = `r sum_is_oa$n[2]` out of `r sum(sum_is_oa$n[1:2])` where the status was known); [@fig-pubsummary]A]; however, `r sum_is_oa$valid_percent[1]*100`% (*n* = `r sum_is_oa$n[1]`) were published behind paywalls. Published articles behind paywalls would otherwise remain inaccessible if it were not for *EcoEvoRxiv*. For preprints and postprints published in open access journals, the type of open access also varied widely (e.g., Gold, Hybrid, Green OA etc., [@fig-pubsummary]B). Such a result may not be too surprising given that authors using preprint servers are probably already ‘pro-open access’; particularly given that *EcoEvoRxiv* is run by Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE).

Data and code sharing are also key components of open science [@roche2015public]. In the spirit of 'openness', we expected data and code sharing among preprints and postprints to be greater than in many papers published in research journals [@roche2015public; @o2021preferred]. Despite this, we found that `r (preprint_data[1,"n"]/sum(preprint_data[c(1,3),"n"]))*100`%  (*n* = `r preprint_data[1,2]`) of articles relying on data (i.e., classified as ‘research articles’) on *EcoEvoRxiv* did not share data, and `r (preprint_code[1,"n"]/sum(preprint_code[c(1,3),"n"]))*100`% (*n* = `r preprint_code[1,2]`) did not share code.

Authors may be reluctant to share data and code for preprints because of the perceived concern that others may acquire and use their data and code before publication in a journal. Authors of `r pre_article_data[3,"percent"]*100`% (*n* = `r pre_article_data[3,2]`) of articles that did not share data at the preprint stage did ultimately share data when the article was published; whereas authors of `r pre_article_data[2,"percent"]*100`% (*n* = `r pre_article_data[2,2]`) never shared data. However, `r pre_article_data[1,"percent"]*100`% (*n* = `r pre_article_data[1,2]`) shared data at both stages. The same was true for code. Overall, `r pre_article_code[3,"percent"]*100`% (*n* = `r pre_article_code[3,2]`) of preprints had no open code at the preprint stage but did at the published article stage and authors of `r pre_article_code[2,"percent"]*100`% (*n* = `r pre_article_code[2,2]`) preprints did not share code at either stage. However, `r pre_article_code[1,"percent"]*100`% (*n* = `r pre_article_code[1,2]`) shared code at both stages. Relatively low code and data-sharing practices in our sample is consistent with analyses of sharing practices for published articles (e.g., [@o2021preferred]), even for journals with strict public data archiving policies [@roche2015public].

@@

Our analysis has allowed us to better understand preprinting/postprinting practices in *EcoEvoRxiv*. Overall, *EcoEvoRxiv* articles are diverse but with primary research articles on vertebrates comprising most of the articles posted. North America, Europe and Australia use *EcoEvoRxiv* the most with very few non-English language articles to date. Submitting authors who were earlier in their career and more often with ‘male-associated names’ tended to use  *EcoEvoRxiv* the most. Articles posted to *EcoEvoRxiv* tend to take up to 8 months to become published with many articles not being open access. Code and data sharing was also relatively uncommon at the preprint stage. At the same time, we attempted to collect data on community discussion around preprints. However, no such data was found on preprint landing pages, likely reflecting inadequate functionality and cross-linking with sources where such discussion is occurring. Based on the insights from our analysis, we provide recommendations to authors and the scientific community on ways they can further promote open and transparent research through preprints:

- First, share your data and code at the preprint stage. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings [@sarabipour2019value; @gomes2022don]. Reassuringly, sharing data and code is rarely associated with the 'scooping' of research findings [@soeharjono2021reported]. If authors are worried about data being used unintentionally, clear information surrounding its reuse can be included in a license (see https://choosealicense.com). Data can also be archived with an embargo on its reuse [@roche2014troubleshooting]. 
